---
title: "Castle Law Replication: Lee & Wooldridge (2025) Section 7.2"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Castle Law Replication}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Overview

This vignette replicates the staggered DiD estimates from Lee & Wooldridge (2025),
Section 7.2, using the `lwdidR` package. The analysis studies the effect of
Castle Doctrine laws on log homicide rates for US states, 2000–2010.

**Expected results (from paper and Python replication):**

| Estimator | ATT | SE | t-stat |
|-----------|-----|-----|--------|
| Demeaning (OLS SE) | ≈ 0.092 | ≈ 0.057 | ≈ 1.61 |
| Demeaning (HC3 SE) | ≈ 0.092 | ≈ 0.061 | ≈ 1.50 |
| Detrending (HC3 SE) | ≈ 0.067 | ≈ 0.055 | ≈ 1.21 |

---

## Setup

```{r load-package}
library(lwdidR)
library(knitr)
```

```{r load-data}
data_path <- system.file("extdata", "castle.csv", package = "lwdidR")
stopifnot(nchar(data_path) > 0)
castle <- read.csv(data_path)

# Construct gvar: first treatment year (NA for never treated)
castle$gvar <- castle$effyear
castle$gvar[is.na(castle$gvar) | castle$gvar == 0] <- NA

# Verify data structure
cat("Dimensions:", nrow(castle), "x", ncol(castle), "\n")
cat("Years:", min(castle$year), "to", max(castle$year), "\n")
cat("States:", length(unique(castle$sid)), "\n")
cat("Treatment cohorts:\n")
cohort_counts <- table(castle$gvar[!duplicated(castle$sid)], useNA = "ifany")
print(cohort_counts)
```

---

## Demeaning Estimator

### With OLS Standard Errors

```{r demean-ols}
res_demean_ols <- lwdid(
  data          = castle,
  y             = "lhomicide",
  ivar          = "sid",
  tvar          = "year",
  gvar          = "gvar",
  rolling       = "demean",
  control_group = "never_treated",
  aggregate     = "overall",
  vce           = NULL  # homoskedastic OLS
)
print(res_demean_ols)
```

### With HC3 Standard Errors

```{r demean-hc3}
res_demean_hc3 <- lwdid(
  data          = castle,
  y             = "lhomicide",
  ivar          = "sid",
  tvar          = "year",
  gvar          = "gvar",
  rolling       = "demean",
  control_group = "never_treated",
  aggregate     = "overall",
  vce           = "hc3"
)
print(res_demean_hc3)
```

---

## Detrending Estimator

```{r detrend-hc3}
res_detrend_hc3 <- lwdid(
  data          = castle,
  y             = "lhomicide",
  ivar          = "sid",
  tvar          = "year",
  gvar          = "gvar",
  rolling       = "detrend",
  control_group = "never_treated",
  aggregate     = "overall",
  vce           = "hc3"
)
print(res_detrend_hc3)
```

---

## Summary Comparison Table

```{r comparison-table}
paper_vals <- data.frame(
  Method       = c("Demeaning (OLS)", "Demeaning (HC3)", "Detrending (HC3)"),
  Paper_ATT    = c(0.092, 0.092, 0.067),
  Paper_SE     = c(0.057, NA, 0.055),
  Paper_tstat  = c(1.61, 1.50, 1.21)
)

r_vals <- data.frame(
  Method     = c("Demeaning (OLS)", "Demeaning (HC3)", "Detrending (HC3)"),
  R_ATT      = round(c(res_demean_ols$att_overall,
                        res_demean_hc3$att_overall,
                        res_detrend_hc3$att_overall), 4),
  R_SE       = round(c(res_demean_ols$se_overall,
                        res_demean_hc3$se_overall,
                        res_detrend_hc3$se_overall), 4),
  R_tstat    = round(c(res_demean_ols$tstat,
                        res_demean_hc3$tstat,
                        res_detrend_hc3$tstat), 3)
)

results_table <- merge(paper_vals, r_vals, by = "Method")
kable(results_table, caption = "Table 7.2 Replication: Lee & Wooldridge (2025)")
```

---

## Cohort-Specific Effects

```{r cohort-effects}
res_cohort <- lwdid(
  data          = castle,
  y             = "lhomicide",
  ivar          = "sid",
  tvar          = "year",
  gvar          = "gvar",
  rolling       = "demean",
  control_group = "never_treated",
  aggregate     = "cohort",
  vce           = "hc3"
)

cat("\nCohort-specific ATT estimates:\n")
kable(
  res_cohort$att_by_cohort[, c("cohort", "att", "se", "ci_lower", "ci_upper",
                                 "tstat", "pvalue", "n_periods", "n_units")],
  digits = 4,
  caption = "Cohort Effects (Equation 7.9, Lee & Wooldridge 2025)"
)
```

---

## All (g, r) Effects

```{r gr-effects}
res_none <- lwdid(
  data          = castle,
  y             = "lhomicide",
  ivar          = "sid",
  tvar          = "year",
  gvar          = "gvar",
  rolling       = "demean",
  control_group = "never_treated",
  aggregate     = "none",
  vce           = "hc3"
)

cat(sprintf("\nTotal (g,r) pairs estimated: %d\n", nrow(res_none$att_by_cohort_time)))
kable(
  head(res_none$att_by_cohort_time[, c("cohort", "period", "event_time",
                                         "att", "se", "pvalue")], 15),
  digits = 4,
  caption = "(g, r)-Specific Effects (Equation 7.8, Lee & Wooldridge 2025)"
)
```

---

## Numerical Checks

```{r checks}
cat("=== Numerical Verification ===\n\n")

tol <- 0.001
checks <- list(
  list(label = "Demeaning ATT",   got = res_demean_ols$att_overall, ref = 0.092),
  list(label = "OLS SE",          got = res_demean_ols$se_overall,  ref = 0.057),
  list(label = "OLS t-stat",      got = res_demean_ols$tstat,       ref = 1.607),
  list(label = "HC3 t-stat",      got = res_demean_hc3$tstat,       ref = 1.50),
  list(label = "Detrending ATT",  got = res_detrend_hc3$att_overall,ref = 0.067),
  list(label = "Detrend HC3 t",   got = res_detrend_hc3$tstat,      ref = 1.21)
)

all_pass <- TRUE
for (chk in checks) {
  diff  <- abs(chk$got - chk$ref)
  pass  <- diff < tol
  all_pass <- all_pass && pass
  cat(sprintf("%-20s  got: %7.4f  ref: %5.3f  diff: %.5f  %s\n",
              chk$label, chk$got, chk$ref, diff,
              if (pass) "PASS" else "FAIL"))
}
cat(sprintf("\nAll checks pass: %s\n", all_pass))
```

---

## Session Info

```{r session}
sessionInfo()
```
