---
title: "Large-N Rolling DiD: The Walmart Application (Lee & Wooldridge 2026a)"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Large-N Rolling DiD: The Walmart Application (Lee & Wooldridge 2026a)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4.2, cache = TRUE)
walmart_path <- system.file("extdata", "lw_walmart.rds", package = "lwdidR")
have_data <- nzchar(walmart_path) && file.exists(walmart_path)
knitr::opts_chunk$set(eval = have_data)
```

## Overview

The **large-N path** of `lwdidR` implements the doubly-robust rolling
difference-in-differences estimator of Lee & Wooldridge (2026a). It is activated
by setting `method` to one of:

- `"ra"` — regression adjustment,
- `"ipw"` — inverse-probability weighting,
- `"ipwra"` — doubly-robust IPWRA (inverse-probability-weighted regression adjustment).

For each treatment cohort `g` and calendar period `t`, the outcome is residualised
by removing each unit's own pre-treatment mean (`rolling = "demean"`) or linear
trend (`rolling = "detrend"`); an ATT(g,t) is then estimated on the eligible
control cross-section. The cell estimates are aggregated to the **event-study
WATT(r)** path (cohort-size-weighted by relative period r, where r = t − g) plus overall
**Pre/Post averages**. Inference uses a **wild cluster bootstrap** with
**simultaneous (sup-t) confidence bands** for the event-study path; the Pre/Post
averages use pointwise normal intervals.

> This vignette uses the `lw_walmart` data from the original Stata `lwdid`
> package, bundled with `lwdidR` (see Citation). It revisits the effect of
> Walmart store openings on county-level employment.

```{r data}
library(lwdidR)
walmart <- readRDS(system.file("extdata", "lw_walmart.rds", package = "lwdidR"))
str(walmart)
```

Variables: `cid` (county id), `year`, `first_year` (cohort; `0` = never opened a
Walmart in-window), `log_wholesale_emp` (outcome), and covariates `x1`, `x2`, `x3`.

## Regression adjustment (demean)

```{r ra}
ra <- lwdid(walmart, "log_wholesale_emp", "cid", "year", gvar = "first_year",
            rolling = "demean", method = "ra",
            controls = c("x1", "x2", "x3"), reps = 199, seed = 1)
print(ra)
```

These estimates reproduce the Stata `lwdid` v2.4 output to ~`1e-9` for the point
estimates, with bootstrap standard errors agreeing to within Monte-Carlo error
(see `data-raw/stata-reference/`).

## Doubly-robust IPWRA (detrend)

This mirrors the recommended specification when parallel trends may be violated:
unit-specific **detrending** with the **doubly-robust** estimator.

```{r ipwra}
dr <- lwdid(walmart, "log_wholesale_emp", "cid", "year", gvar = "first_year",
            rolling = "detrend", method = "ipwra",
            controls = c("x1", "x2", "x3"), reps = 199, seed = 1)
print(dr)
```

### Event-study plot

```{r plot}
plot(dr)
```

The bars are **simultaneous** (sup-t) confidence bands across event times; the
pre-treatment periods provide an informal parallel-trends check (with the
detrending normalisation anchoring r = −1 and r = −2 to zero).

## Options

| Option | Meaning |
|---|---|
| `method` | `"ra"`, `"ipw"`, `"ipwra"` (required for the large-N path) |
| `rolling` | `"demean"` or `"detrend"` |
| `controls` | covariate names (required for `ipw`/`ipwra`) |
| `pre` | number of most-recent pre-periods used in the transform (`-1` = all) |
| `never` | use never-treated controls only (default: not-yet-treated too) |
| `attgt` | also return the ATT(g,t) cell estimates |
| `ydot` | also return the per-cohort transformed outcomes |
| `reps`, `seed` | wild cluster bootstrap replications and seed |
| `level`, `nose` | confidence level; skip SEs |

For production use, increase `reps` (e.g., 999) for stable bootstrap inference.

## Citation

`lwdidR` is an R port of the original Stata `lwdid` command by Soo Jeong Lee and
Jeffrey M. Wooldridge (<https://github.com/Soo-econ/lwdid>); the bundled
`lw_walmart` data ships with that package. The large-N method is:

> Lee, S. J., & Wooldridge, J. M. (2026a). Simple Transformation Approach to
> Difference-in-Differences Estimation for Panel Data. *Journal of Business &
> Economic Statistics*, 1–27. <https://doi.org/10.1080/07350015.2026.2683047>
```
